\Large Clustering Vehicles based on Trips Identified from Automatic \\ Number Plate Recognition Camera Scans

### 2018-02-23 dataset.tex
Made some text changes and some re-placement of tables/figures: discussion/errors in plate scanning
Latest changes with suggested fixes by supervisors.

\begin{table}[!ht]
\centering
\tabcolsep=0.17cm
\small
\begin{tabular}{c c c c c}
  \hline
  Vehicle & Camera & Timestamp & \thead{Clock \\Error} & Confidence \\
  \hline
  169239 & 1031 & 1454284800.26 &   0 & 100 \\
  12862943 & 18 & 1454284800.97 &   8 &  61 \\
  16243894 & 22 & 1454284801.46 &   6 &  86 \\
  4817789 & 52 & 1454284803.43 &  13 &  94 \\
  5503486 & 110 & 1454284802.19 &  22 &  91 \\
  15244177 & 115 & 1454284802.83 &  18 &  87 \\
   \hline
\end{tabular}
\caption{Sample of number plate data. Clock error is given in milliseconds and confidence as a percentile value.}
\label{t:np_data_example}
%\vspace{-2cm}
\end{table}

@@

\begin{figure*}[t]
\centering
\begin{subfigure}[t]{.48\textwidth}
  \centering
  \includegraphics[width=.9\linewidth]{observations_per_day.pdf}
  \caption{Total number of scans recorded per day in Tyne and Wear. There is a clear seasonal effect caused by decreasing traffic demands at weekends and increasing traffic volume during weekdays.}
  \label{fig:observations-per-day}
\end{subfigure}\hfill
\begin{subfigure}[t]{.48\textwidth}
  \centering
  \includegraphics[width=.9\linewidth]{observations_per_day_camera.pdf}
  \caption{Number of scans recorded per ANPR camera and day in Tyne and Wear. Inter-camera variability is observed, as some cameras are located in more traffic intensive road sections than others. Decommissioned or temporarily unavailable cameras (due to loss of power, faulty camera, road closed, etc) can be identified at the bottom.}
  \label{fig:observations-per-camera-day}
\end{subfigure}
\caption{License plate scans recorded by ANPR cameras during February 2017, in the region of Tyne and Wear, United Kingdom.}
\label{fig:time-series}
\end{figure*}

In the region of Tyne and Wear, United Kingdom, there are over 250 active ANPR cameras. Over 1 million license plate detections are recorded by these cameras every day. Figure~\ref{fig:time-series} shows the number of daily scans recorded over a month (February, 2017). Furthermore, every scan is stored in a central database managed by Urban Traffic Management and Control (UTMC) Tyne and Wear, and used to compute travel times across particular links of interest in the road network. These are usually major roads that see high volumes of traffic, or road segments more prone to traffic jams. Average journey times can then be conveyed back to the drivers by the way of Variable Message Signs (VMS) or web based applications. Figure~\ref{fig:anpr-overview} represents this interaction.

Number plate data, in its essence, is a stream of events, each representing a vehicle observed by one camera at a specific point in time. An excerpt of the data can be found in table~\ref{t:np_data_example}. All number plate were anonymised by the UTMC through a hashing algorithm before the data was shared. Cameras are uniquely identified by an integer and timestamps are relative to each camera's clock. Clock synchronisation is performed using the Network Time Protocol (NTP), as the cameras are connected, through a private network, to a central server. Therefore, the recorded timestamps can be used directly if the synchronisation error is negligible. The following additional information is also captured and provided by each camera:
\begin{enumerate*}[label=(\roman*)]
  \item the clock synchronisation error (milliseconds);
  \item the camera's confidence that the identified number plate is the true number plate (percentage);
  \item the direction of travel, away or towards the camera.
\end{enumerate*}
The confidence in the observation is especially useful as it helps diagnosing license plate recognition errors. On the other hand, the direction of travel is dependent upon the orientation of the camera, which is not provided. Hence, we chose to ignore the latter in this work.

>>>>

\begin{figure}[t]
\centering
\includegraphics[width=.90\linewidth]{ANPR-overview.pdf}
\caption{Overview of a ANPR-based system for traffic monitoring and control.}
\label{fig:anpr-overview}
\vspace{-0.25cm}
\end{figure}

\begin{figure*}[!ht]
\centering
\begin{subfigure}[t]{.48\textwidth}
  \centering
  \includegraphics[width=.9\linewidth]{observations_per_day.pdf}
  \caption{Total number of scans recorded per day in Tyne and Wear. There is a clear seasonal effect caused by decreasing traffic demands at weekends and increasing traffic volume during weekdays.}
  \label{fig:observations-per-day}
\end{subfigure}\hfill
\begin{subfigure}[t]{.48\textwidth}
  \centering
  \includegraphics[width=.9\linewidth]{observations_per_day_camera.pdf}
  \caption{Number of scans recorded per ANPR camera and day in Tyne and Wear. Inter-camera variability is observed, as some cameras are located in more traffic intensive road sections than others. Decommissioned or temporarily unavailable cameras (due to loss of power, faulty camera, road closed, etc) can be identified at the bottom.}
  \label{fig:observations-per-camera-day}
\end{subfigure}
\caption{License plate scans recorded by ANPR cameras during February 2017, in the region of Tyne and Wear, United Kingdom.}
\label{fig:time-series}
\vspace{-0.4cm}
\end{figure*}

@@

In the region of Tyne and Wear, United Kingdom, there are over 250 active ANPR cameras. Over 1 million license plate detections are recorded by these cameras every day. Figure~\ref{fig:time-series} shows the number of daily scans recorded over a month (February, 2017). Furthermore, every scan is stored in a central database managed by the Tyne and Wear UTMC, and used to compute travel times across particular links of interest in the road network. These are usually major roads that see high volumes of traffic, or road segments more prone to traffic jams. Average journey times can then be conveyed back to the drivers by the way of Variable Message Signs (VMS) or web based applications. Figure~\ref{fig:anpr-overview} represents this interaction.

Number plate data, in its essence, is a stream of events, each representing a vehicle observed by one camera at a specific point in time. An excerpt of the data can be found in Table~\ref{t:np_data_example}. All number plate were anonymised by the UTMC through a hashing algorithm before the data was shared. Cameras are uniquely identified by an integer and timestamps are relative to each camera's clock. Clock synchronisation is performed using the Network Time Protocol (NTP), as the cameras are connected, through a private network, to a central server. Therefore, the recorded timestamps can be used directly if the synchronisation error is negligible. The following additional information is also captured and provided by each camera:
\begin{enumerate*}[label=(\roman*)]
  \item the clock synchronisation error (milliseconds);
  \item the camera's confidence that the identified number plate is the true number plate (percentage);
  \item the direction of travel, away or towards the camera.
\end{enumerate*}
The confidence in the observation is especially useful as it helps diagnosing license plate recognition errors. On the other hand, the direction of travel is dependent upon the orientation of the camera, which is not provided. Hence, we chose to ignore the latter in this work.

\begin{table}[t]
\centering
\tabcolsep=0.12cm
\begin{tabular}{c c c c c}
  \hline
  Vehicle & Camera & Timestamp & \thead{Clock \\Error} & Confidence \\
  \hline
  169239 & 1031 & 1454284800.26 &   0 & 100 \\
  12862943 & 18 & 1454284800.97 &   8 &  61 \\
  16243894 & 22 & 1454284801.46 &   6 &  86 \\
  4817789 & 52 & 1454284803.43 &  13 &  94 \\
  5503486 & 110 & 1454284802.19 &  22 &  91 \\
  15244177 & 115 & 1454284802.83 &  18 &  87 \\
   \hline
\end{tabular}
\caption{Sample of number plate data. Clock error is given in milliseconds and confidence as a percentile value.}
\label{t:np_data_example}
\vspace{-0.3cm}
\end{table}

### 2018-02-23 discussion.tex
Latest changes with suggested fixes by supervisors.

Most urban cities in the world employ a network of ANPR cameras that are used for law enforcement as well as traffic monitoring and control. Number plate data collected in the Tyne and Wear area is stored and leveraged by the UTMC for computing average journey times across a selection of sensitive roads. However, number plate data could be used more extensively to identify and study individual and collective travel patterns. In this paper, we have presented a set of definitions and constraints that establish a conceptual foundation for identifying vehicle trips from number plate detections.

@@

Once trip data has been computed, a range of interesting applications are available. This work tries to identify groups of vehicles by clustering information about frequency and diversity of travel. By associating a vehicle with a cluster that represents taxis, or home-to-work commuters, one can begin estimating trip mode usage across the city. However, the results presented here could benefit from extra work and further validation. Namely, gaining access to a database that maps plate numbers to vehicle types would not only provide means to better validate the proposed methodology but also enable the application of supervised learning instead. Furthermore, this work can be improved upon by using other data sources as a baseline for evaluating the reliability of extracted trips, such as GPS taxi traces. Finally, future work can focus on using trip data to solve interesting research problems such as:
\begin{enumerate*}[label=(\roman*)]
  \item real-time route recommendation using probabilistic graphical models;
  \item detection of abnormal trip patterns for helping law enforcement in the identification of suspect vehicles or behaviour;
  \item modelling how drivers make routing choices in the presence of anomalies in the road network.
\end{enumerate*}

\subsection*{Acknowledgments}

>>>>

Most urban cities in the world employ a network of ANPR cameras that are used for law enforcement as well as traffic monitoring and control. Number plate data collected in the Tyne and Wear area is stored and leveraged by the Urban Traffic and Management Centre (UTMC) for computing average journey times across a selection of sensitive roads. However, number plate data could be used more extensively to identify and study individual and collective travel patterns. In this paper, we have presented a set of definitions and constraints that establish a conceptual foundation for identifying vehicle trips from number plate detections.

@@

Once trip data has been computed, a range of interesting applications are available. This work tries to identify groups of vehicles by clustering information about frequency and diversity of travel. By associating a vehicle with a cluster that represents taxis, or home-to-work commuters, one can begin estimating trip mode usage across the city. However, the results presented here could benefit from extra work and further validation. Namely, gaining access to a database that maps plate numbers to vehicle types would not only provide means to better validate the proposed methodology but also enable the application of supervised learning instead. Furthermore, this work can be improved upon by using other data sources as a baseline for evaluating the reliability of extracted trips, such as GPS taxi traces.

Finally, future work can focus on using trip data to solve interesting research problems such as:
\begin{enumerate*}[label=(\roman*)]
  \item real-time route recommendation using probabilistic graphical models;
  \item detection of abnormal trip patterns for helping law enforcement in the identification of suspect vehicles or behaviour;
  \item modelling how drivers make routing choices in the presence of anomalies in the road network.
\end{enumerate*}

\vspace{-0.3cm}
\subsection*{Acknowledgments}

### 2018-02-23 paper.tex
Latest changes with suggested fixes by supervisors.

\documentclass[twoside,leqno,twocolumn]{article}
\usepackage{ltexpprt}
\usepackage{parskip}
\usepackage[inline]{enumitem}
\usepackage{amsmath}
\usepackage[many]{tcolorbox}
\usepackage{graphicx}
\usepackage{subcaption}
\usepackage{makecell}
\usepackage{float}
\usepackage{placeins}
\usepackage{microtype}

@@

\title{\Large Clustering Vehicles based on Trips Identified from Automatic \\ Number Plate Recognition Camera Scans}
\author{Pedro M. Pinto Silva \thanks{School of Computing, Newcastle University, United Kingdom}
\and
Matthew Forshaw\footnotemark[1]
\and
A. Stephen McGough\footnotemark[1]}
\date{}

@@

% Default Copyright Statement
\fancyfoot[R]{\footnotesize{\textbf{Copyright \textcopyright\ 2018 by BigTraffic\\
Unauthorized reproduction of this article is prohibited}}}

@@

\begin{abstract} \small\baselineskip=9pt In major cities, government agencies increasingly employ automatic number-plate recognition (ANPR) technology in law enforcement and traffic control. In the Tyne and Wear region (UK) the network of ANPR cameras is used to monitor travel times across sensitive roads. So far, few works have explored the full potential of number-plate scans for analysing individual and collective travel patterns. In this work we present a methodology for deriving trips from vehicle sightings at fixed camera locations. We illustrate the effect of parameters $\tau$ and $\Tau$ on trip discrimination and on the detection of implausible trips. To demonstrate the potential of trip data we apply \emph{k-means} clustering to trips identified from over 40 million plate scans recorded over fifteen weekdays. Results show that whilst private and transit travel modes can begin to be inferred from the resulting clusters, further work needs to be put into developing a more consistent and integrated framework for trip identification in ANPR data.\end{abstract}

@@

\bibitem{ODMobileData}
Alexander, Lauren, et al. "Origin–-destination trips by purpose and time of day inferred from mobile phone data." Transportation research part c: emerging technologies 58 (2015): 240-250.

\bibitem{Castillo2010}
Castillo, Enrique, et al. {\em Optimal use of plate-scanning resources for route flow estimation in traffic networks}. IEEE Transactions on Intelligent Transportation Systems 11.2 (2010): 380-391.

\bibitem{Castillo2008}
Castillo, Enrique, José María Menéndez, and Pilar Jiménez. {\em Trip matrix and path flow reconstruction and estimation based on plate scanning and link observations}. Transportation Research Part B: Methodological 42.5 (2008): 455-481.

\bibitem{Clustering}
Chen, Huiyu, Chao Yang, and Xiangdong Xu. {\em Clustering Vehicle Temporal and Spatial Travel Behavior Using License Plate Recognition Data}. Journal of Advanced Transportation 2017 (2017).

@@

\bibitem{EvolutionUTMC2013}
Hamilton, Andrew, et al. {\em The evolution of urban traffic control: changing policy and technology}. Transportation planning and technology 36.1 (2013): 24-43.

@@

\bibitem{Hazelton2012}
Parry, Katharina, and Martin L. Hazelton. {\em Estimation of origin-–destination matrices from link counts and sporadic routing data}. Transportation Research Part B: Methodological 46.1 (2012): 175-188.

\bibitem{ClusteringGPS}
Schüssler, Nadine, and Kay W. Axhausen. {\em Identifying trips and activities and their characteristics from GPS raw data without further information}. Arbeitsbericht Verkehrs-und Raumplanung 502 (2008).

\bibitem{Kmeans}
Steinley, Douglas, and Michael J. Brusco. {\em Choosing the number of clusters in Κ-means clustering}. Psychological Methods 16.3 (2011): 285.

\bibitem{SurveyITS2011}
Zhang, Junping, et al. {\em Data-driven intelligent transportation systems: A survey}. IEEE Transactions on Intelligent Transportation Systems 12.4 (2011): 1624-1639.

>>>>

\documentclass[twoside,leqno,twocolumn]{article}
\usepackage{ltexpprt}
\usepackage{parskip}
\usepackage[inline]{enumitem}
\usepackage{amsmath}
\usepackage[many]{tcolorbox}
\usepackage{graphicx}
\usepackage{subcaption}
\usepackage{makecell}
\usepackage{float}
\usepackage{placeins}
\usepackage{microtype}
\usepackage{enumitem}

@@

\title{\Large Clustering Vehicles based on Trips Identified from Automatic \\ Number Plate Recognition Camera Scans}
\author{Pedro M. Pinto Silva \thanks{School of Computing, Newcastle University, United Kingdom\break \{p.pinto-da-silva2, matthew.forshaw, stephen.mcgough\}@newcas\-tle.ac.uk}
\and
Matthew Forshaw\footnotemark[1]
\and
A. Stephen McGough\footnotemark[1]}
\date{}

@@

% Default Copyright Statement
\fancyfoot[R]{\footnotesize{\textbf{Copyright \textcopyright\ 2018\\
Unauthorized reproduction of this article is prohibited}}}

@@

\begin{abstract} \small\baselineskip=9pt In major cities, government agencies increasingly employ automatic number-plate recognition (ANPR) technology in law enforcement and traffic control. In the Tyne and Wear region (UK) the network of ANPR cameras is used to monitor travel times across sensitive roads. So far, few works have explored the full potential of number-plate scans for analysing individual and collective travel patterns. In this work we present a methodology for deriving trips from vehicle sightings at fixed camera locations. We identify two parameters $\tau$ and $\Tau$ as essential for identifying implausible trips and differentiating between multiple trips of the same vehicle. To demonstrate the applicability of trip data we apply \emph{k-means} clustering to trips identified from over 40 million plate scans recorded over fifteen weekdays. Results show that whilst private and transit travel modes can begin to be inferred from the resulting clusters, further work needs to be put into developing a more consistent and integrated framework for trip identification in ANPR data.\end{abstract}

@@

\bibitem{ODMobileData}
Alexander, L., Jiang, S., Murga, M., \& González, M. C. "Origin–-destination trips by purpose and time of day inferred from mobile phone data." Transportation research part c: emerging technologies 58 (2015): 240-250.

\bibitem{Castillo2010}
Castillo, E. Gallego, I., Menéndez, J. M., \& Rivas, A {\em Optimal use of plate-scanning resources for route flow estimation in traffic networks}. IEEE Transactions on Intelligent Transportation Systems 11.2 (2010): 380-391.

\bibitem{Castillo2008}
Castillo, E., Menéndez, J. M., \& Jiménez, P. {\em Trip matrix and path flow reconstruction and estimation based on plate scanning and link observations}. Transportation Research Part B: Methodological 42.5 (2008): 455-481.

\bibitem{Clustering}
Chen, H., Yang, C., \& Xu, X. {\em Clustering Vehicle Temporal and Spatial Travel Behavior Using License Plate Recognition Data}. Journal of Advanced Transportation 2017 (2017).

@@

\bibitem{EvolutionUTMC2013}
Hamilton, Andrew, Waterson, B., Cherrett, T., Robinson, A., \& Snell, I. {\em The evolution of urban traffic control: changing policy and technology}. Transportation planning and technology 36.1 (2013): 24-43.

@@

\bibitem{Hazelton2012}
Parry, K., \& Hazelton, M. L. {\em Estimation of origin-–destination matrices from link counts and sporadic routing data}. Transportation Research Part B: Methodological 46.1 (2012): 175-188.

\bibitem{ClusteringGPS}
Schüssler, N., \& Axhausen, K. W. {\em Identifying trips and activities and their characteristics from GPS raw data without further information}. Arbeitsbericht Verkehrs-und Raumplanung 502 (2008).

\bibitem{Kmeans}
Steinley, D., \& Brusco, M. J. {\em Choosing the number of clusters in Κ-means clustering}. Psychological Methods 16.3 (2011): 285.

\bibitem{SurveyITS2011}
Zhang, J., Wang, F. Y., Wang, K., Lin, W. H., Xu, X., \& Chen, C. {\em Data-driven intelligent transportation systems: A survey}. IEEE Transactions on Intelligent Transportation Systems 12.4 (2011): 1624-1639.

### 2018-02-23 trips.tex
A few changes - submitted to supervisor for review
Latest changes with suggested fixes by supervisors.

% !TEX root = paper.tex
\subsection{Trip Identification}\label{s.trips}

@@

Let an ordered sequence of sightings of vehicle \emph{k} define the $u_{th}$ trip of vehicle \emph{k}:
\begin{equation} \label{e.trip}
w^{k}_{u} = \left(s^{k}_{(1)}, s^{k}_{(2)}, \dots , s^{k}_{(n)}\right),
\end{equation}
where \( n \) is the degree of the trip, i.e.\ the number of sightings. Moreover, let the corresponding journey time sequence, of degree \(n-1\), be defined as the time difference of consecutive sightings:
\begin{equation} \label{e.journeytime}
jt^{k}_{u} = \left(t^{k}_{(2)} - t^{k}_{(1)}, \ldots, t^{k}_{(n)} - t^{k}_{(n-1)} \right).
\end{equation}
We consider a trip of \emph{$w^k_u$} valid under the following conditions:
\begin{align}
n &\ge 1 , \label{e.trip.constraints.1} \\
\tau_{(i)} &< jt^{k}_{u(i)} < \Tau_{(i)} \ , \ \forall i \in jt^{k}_{u(i)} \label{e.trip.constraints.2},
\end{align}
where $\tau_{(i)}$ and $\Tau_{(i)}$ are the lower and upper bound of the $i_{th}$ element of the journey time sequence.

The first condition~\ref{e.trip.constraints.1} is straightforward and specifies that every identified trip should have at least one sighting. Obviously, vehicles can make trips that do not pass through any ANPR cameras and thus have no associated sightings: $n = 0$. However, this work focus on trips that we can observe and hence we consider that $n > 0$. The second condition~\ref{e.trip.constraints.2} defines a minimum and maximum travel times between consecutive observations. Its purpose is twofold:
\begin{enumerate*}[label=(\roman*)]
  \item first, to allow distinct trips made by the same vehicle to be differentiated. For instance, given two consecutive sightings of \emph{k} three hours apart, we likely want to interpret them as belonging to different trips of \emph{k};
  \item second, it allows implausible trips to be identified. For example, an implausible trip can result from observing \emph{k} at a given camera and then a few seconds later at a second camera, several miles apart. Two explanations are common, either one of the cameras made a detection error, or there is another vehicle with a cloned number-plate travelling on the road network.
\end{enumerate*} Evidently, condition~\ref{e.trip.constraints.2} is only valid for trips of degree two sightings or greater. Nevertheless, trips can easily be differentiated by first sorting sightings by time of occurrence, then calculating the journey time sequence for the entire sequence and finally comparing each element against $\Tau$. An example of a trip identified this way can be seen in Figure~\ref{fig:trip-example}.

The simplest approach to choosing the value of $\Tau$ is to pick a fixed empirical value, such as 5 or 10 minutes. However, if the distance between two cameras is greater than another pair of cameras, then it makes sense that $\Tau$ is relaxed. Similarly, if there is an anomaly in the road network, such as a traffic jam, and the routes connecting the two cameras are affected, then the value of $\Tau$ should also be adapted. Hence, $\Tau$ should be a function of the distance between the two cameras (or, more accurately, of the top n-routes between these) and the distribution of observed journey times. The same rationale can be applied to $\tau$. However, due to time constraints, in this work we simply used several fixed values of $\Tau$ and set $\tau$ to zero. Figure~\ref{fig:length-dist} shows how the number of trips per degree of trip varies by fixing $\Tau$ at different empirical values.

@@

\begin{figure*}[!ht]%
  \centering
  \begin{subfigure}[c]{.5\textwidth}
    \small
    \tabcolsep=0.09cm
    \begin{tabular}{c c c c c c c}
      \hline
      Vehicle & Camera & Timestamp & Trip & Sighting & \thead{Journey \\Time} & \thead{Trip \\Id}\\
      \hline
      2362920 & 1014 & \makecell{2017-02-01 \\ 00:00:06} &   1 &   1 & NA & 21 \\
      2362920 & 1044 & \makecell{2017-02-01 \\ 00:01:28} &   1 &   2 & 82.38 & 21 \\
      2362920 &  35 & \makecell{2017-02-01 \\ 00:02:32} &   1 &   3 & 63.50 & 21\\
      2362920 &  32 & \makecell{2017-02-01 \\ 00:04:38} &   1 &   4 & 125.95 & 21\\
       \hline
    \end{tabular}
    \label{fig:trip-example-table}
  \end{subfigure}\hfill
  %\qquad
  \begin{subfigure}[c]{.48\textwidth}
    \centering
    \includegraphics[width=1\linewidth]{trip-example.pdf}
    \label{fig:trip-example-map}
  \end{subfigure}\hfill
  \caption{Example of a trip of degree 4. On the left side the corresponding data table is shown. The $u_{th}$ trip of each vehicle is given by the variable \emph{Trip}, whereas the $i_{th}$ sighting is given by variable \emph{Sighting}. The variable \emph{JourneyTime} gives the travel time from the previous to current sighting. Lastly, the variable \emph{TripId} represents the unique sequence of cameras that describes the trip. This allows trips to be grouped and summarised not only in terms of their origins and destination, but also routes. On the right side, the same trip is plotted on a map. The circles represent camera locations whereas the lines represent the fastest driving routes between sightings rather than the true route taken by the vehicle. Even though no routing information is available for each consecutive pair of sightings, the observed journey times can be compared against the distribution of collective journey times to rank the set of most likely routes chosen (which can be determined for instance by Stochastic User Equilibrium~\cite{Castillo2008}).}%
  \label{fig:trip-example}%
  \vspace{-0.40cm}
\end{figure*}
\vspace{-0.25cm}

\subsection{Duplicate scannings}

@@

 Although the estimation of $\gamma$ carries similar considerations and consequences as those of estimating $\Tau$ and $\tau$, most duplicates can be identified in consecutive sightings of the same camera within the same trip. Even though a poor estimation of $\gamma$ also has an impact on error propagation, this decreases substantially after filtering duplicates according to the heuristic above, due to the low occurrence of cycles in trips.
\vspace{-0.25cm}

\subsection{Errors in plate scanning}

ANPR cameras have average accuracy rates of 99.9\% and higher. If we consider that on average 1 to 10 out of 10000 number plate scans are misclassified number-plates then approximately between 200 to 2000 scans everyday are incorrect. These errors propagate and lead to innacuracies in the identification of trips. In fact, misclassifications affect the trip sequences of two vehicles: the true passing vehicle and the vehicle erroneously detected instead. The true vehicle will be missing a sighting in the correspoding trip sequence vector whereas the other vehicle's trip sequence will contain an extra invalid sighting. Moreover, ANPR cameras may exceptionally fail to detect passing vehicles. Even though only the passing vehicle is affected in this case, it's trip sequence vector is nonetheless affected. Therefore, due to their impact in trip identification, it is important to detect and address missing and misclassified scans.

We filer invalid sightings by removing all sightings containing a \emph{confidence} value below $85\%$. However, we do not address missing scans in this work.

\begin{table}[!ht]
\centering
\tabcolsep=0.25cm
\begin{tabular}{c c c c}
  \hline
Tau & Trips & \thead{Average\\Degree} & \thead{Proportion of Trips \\ Degree 1} \\
  \hline
5 min & 13,603,759 & 1.46 & 0.70 \\
7.5 min & 12,394,709 & 1.60 & 0.64 \\
10 min & 11,690,791 & 1.69 & 0.61 \\
15 min & 10,823,333 & 1.83 & 0.57 \\
20 min & 10,305,860 & 1.92 & 0.54 \\
30 min &  9,653,499 & 2.04 & 0.52 \\
   \hline
\end{tabular}
\caption{Overview of trip data for varying values of $\Tau$.}
\label{t:trips-tau}
\vspace{-0.5cm}
\end{table}

>>>>

% !TEX root = paper.tex
\subsection{Trip identification}\label{s.trips}

@@

Let an ordered sequence of sightings of vehicle \emph{k} define the $u_{th}$ trip of vehicle \emph{k}:
\begin{equation} \label{e.trip}
w^{k}_{u} = \left(s^{k}_{(1)}, s^{k}_{(2)}, \dots , s^{k}_{(n)}\right),
\end{equation}
where \( n \) is the degree of the trip, i.e.\ the number of sightings. Moreover, let the corresponding journey time sequence, of degree \(n-1\), be defined as the time difference of consecutive sightings:
\begin{equation} \label{e.journeytime}
jt^{k}_{u} = \left(t^{k}_{(2)} - t^{k}_{(1)}, \ldots, t^{k}_{(n)} - t^{k}_{(n-1)} \right).
\end{equation}
We consider a trip of \emph{$w^k_u$} valid under the following conditions:
\begin{align}
n &\ge 1 , \label{e.trip.constraints.1} \\
\tau &< jt^{k}_{u(i)} < \Tau \ , \ \forall i \in \{0, 1 ,\ldots, n-1\} \label{e.trip.constraints.2},
\end{align}
where $\tau$ and $\Tau$ are the lower and upper bound of the $i_{th}$ element of the journey time sequence.

\textit{Condition~\ref{e.trip.constraints.1}} is straightforward and specifies that every identified trip should have at least one sighting. Obviously, vehicles can make trips that do not pass through any ANPR cameras and thus have no associated sightings: $n = 0$. However, this work focus on trips that we can observe and hence we consider that $n > 0$.

\textit{Condition~\ref{e.trip.constraints.2}} defines a minimum and maximum travel times between consecutive observations. Its purpose is twofold:
\begin{enumerate*}[label=(\roman*)]
  \item first, to allow distinct trips made by the same vehicle to be differentiated. For instance, given two consecutive sightings of \emph{k} three hours apart, we are likely to want to interpret them as belonging to different trips of \emph{k};
  \item second, it allows implausible trips to be identified. For example, an implausible trip can result from observing \emph{k} at a given camera and then a few seconds later at a second camera, several miles apart. Two explanations are common, either one of the cameras made a detection error, or there is another vehicle with a cloned number-plate travelling on the road network.
\end{enumerate*} Evidently, Condition~\ref{e.trip.constraints.2} is only valid for trips of degree two sightings or greater. Nevertheless, trips can easily be differentiated by first sorting sightings by time of occurrence, then calculating the journey time sequence for the entire sequence and finally comparing each element against $\Tau$. An example of a trip identified this way can be seen in Figure~\ref{fig:trip-example}.

The simplest approach to choosing the value of $\Tau$ is to pick a fixed empirical value, such as 5 or 10 minutes. However, if the distance between two cameras is greater than another pair of cameras, then it makes sense that $\Tau$ is relaxed. Similarly, if there is an anomaly in the road network, such as a traffic jam, and the routes connecting the two cameras are affected, then the value of $\Tau$ should also be adapted. Hence, $\Tau$ should be a function of the distance between the two cameras (or, more accurately, of the top n-routes between these) and the distribution of observed journey times. The same rationale can be applied to $\tau$. However, we focus on $\Tau$ and leave $\tau$ for analysis at a later stage. Figure~\ref{fig:length-dist} shows how the number of trips per degree of trip varies by fixing $\Tau$ at different empirical values.

@@

\begin{figure*}[!ht]%
  \centering
  \begin{subfigure}[c]{.5\textwidth}
    \small
    \tabcolsep=0.09cm
    \begin{tabular}{c c c c c c c}
      \hline
      Vehicle & Camera & Timestamp & Trip & Sighting & \thead{Journey \\Time} & \thead{Trip \\Id}\\
      \hline
      2362920 & 1014 & \makecell{2017-02-01 \\ 00:00:06} &   1 &   1 & NA & 21 \\
      2362920 & 1044 & \makecell{2017-02-01 \\ 00:01:28} &   1 &   2 & 82.38 & 21 \\
      2362920 &  35 & \makecell{2017-02-01 \\ 00:02:32} &   1 &   3 & 63.50 & 21\\
      2362920 &  32 & \makecell{2017-02-01 \\ 00:04:38} &   1 &   4 & 125.95 & 21\\
       \hline
    \end{tabular}
    \label{fig:trip-example-table}
  \end{subfigure}\hfill
  %\qquad
  \begin{subfigure}[c]{.48\textwidth}
    \centering
    \includegraphics[width=1\linewidth]{trip-example.pdf}
    \label{fig:trip-example-map}
  \end{subfigure}\hfill
  \caption{Example of a trip of degree 4. On the left side the corresponding data table is shown. The $u_{th}$ trip of each vehicle is given by the variable \emph{Trip}, whereas the $i_{th}$ sighting is given by variable \emph{Sighting}. The variable \emph{JourneyTime} gives the travel time from the previous to current sighting. Lastly, the variable \emph{TripId} represents the unique sequence of cameras that describes the trip. This allows trips to be grouped and summarised not only in terms of their origins and destination, but also routes. On the right side, the same trip is plotted on a map. The circles represent camera locations whereas the lines represent the fastest driving routes between sightings rather than the true route taken by the vehicle. Even though no routing information is available for each consecutive pair of sightings, the observed journey times can be compared against the distribution of collective journey times to rank the set of most likely routes chosen (which can be determined for instance by Stochastic User Equilibrium~\cite{Castillo2008}).}%
  \label{fig:trip-example}%
  \vspace{-0.40cm}
\end{figure*}

\subsection{Duplicate scannings}\label{s.duplicates}

@@

 Although the estimation of $\gamma$ carries similar considerations and consequences as those of estimating $\Tau$ and $\tau$, most duplicates can be identified in consecutive sightings of the same camera within the same trip. Even though a poor estimation of $\gamma$ also has an impact on error propagation, this decreases substantially after filtering duplicates according to the heuristic above, due to the low occurrence of cycles in trips.

\subsection{Errors in plate scanning}\label{s.errors}

ANPR cameras have an average accuracy rate of 99.9\% or higher. If we consider that on average 1 to 10 out of 10000 number plate scans are misclassified number-plates then approximately between 200 to 2000 scans everyday are incorrect. These errors propagate and lead to innacuracies in the identification of trips. In fact, misclassifications affect the trip sequences of two vehicles: the true passing vehicle and the vehicle erroneously detected instead. The true vehicle will be missing a sighting in the correspoding trip sequence vector whereas the other vehicle's trip sequence will contain an extra invalid sighting. The later may be more easily detected and removed than the former as it may generate a sighting for a vehicle that is normally seen in a different part of the country.

\begin{table*}[t]
\centering
\begin{tabular}{c c c c c c c c c c c}
  \hline
 \thead{Total\\Trips} & \thead{Average\\Trips} & \thead{Average\\Degree} & \thead{Average\\Sightings} & \thead{Average\\Distinct\\Origins} & \thead{Average\\Distinct\\Destinations} & \thead{Average\\Distinct\\Routes} & \thead{Average\\First\\Hour} & \thead{Average\\Last\\Hour} & \thead{Average\\Hour\\Difference} & \thead{Average\\Rest\\Time} \\
  \hline
41 & 3.42 & 1.25 & 5.75 & 2.75 & 1.00 & 3.00 & 15.33 & 19.25 & 3.80 & 3.70 \\
3 & 1.50 & 1.00 & 1.50 & 1.50 & 0.00 & 1.50 & 14.00 & 14.50 & 0.73 & 0.73 \\
7 & 2.33 & 1.33 & 3.33 & 2.33 & 0.67 & 2.33 & 11.00 & 13.00 & 2.44 & 2.41 \\
12 & 2.40 & 1.10 & 3.60 & 2.40 & 0.60 & 2.40 & 14.40 & 16.40 & 2.00 & 1.95 \\
   \hline
\end{tabular}
\caption{Sample of extracted features from trips taken from 15 weekdays of number plate data.}
\vspace{-0.2cm}
\label{t:features}
\end{table*}

Moreover, ANPR cameras may exceptionally fail to detect passing vehicles. Even though only the passing vehicle is affected in this case, it's trip sequence vector is nonetheless affected. Therefore, due to their impact in trip identification, it is important to detect and address missing and misclassified scans. We filter invalid sightings by removing all sightings containing a \emph{confidence} value below $85\%$. However, we do not address missing scans in this work.
